Slovene NLTK Tagger

### 2012-01-08 paper/paper.tex
Whole MSD in table

We must exmplain, that here only the part of whole MSD is printed. For example the whole MSD for \textit{Lep} is:\\\\
\textit{pridevnik vrsta=splošni spol=moški število=ednina sklon=imenovalnik živost=0 vid=0 oblika=0 oseba=0 nikalnost=0 stopnja=nedoločeno določnost=ne število\_svojine=0 spol\_svojine=0 naslonskost=0 zapis=0}\\
\par
The first output was creted with function \textit{prettyPrintWithDescription}, full output can be created with \textit{prettyPrintWithFullDescription}. Both function are defined in \texttt{example.py}
\par
But there is an issue. Because Slovenian language contain letters with caron, the output will have unicode symbols for them. In the previous examples, the output is already corrected. Solution is described in the following section. 
\subsection{Encoding problems} %BANIČ

Printing the result of our example is not optimal. Because we want users to know what these characters are, we implemented a solution. To resolve this, we must decode every word to \textit{UTF - 8} (because words we tag are in UTF - 8). If a user wants to print the result in "pretty" form, he should call function \textit{prettyPrint()} or \textit{prettyPrintByLine()} (both are implemented in \texttt{example.py}).

@@

\section{Conclusion} %BANIČ
To conclude our paper, tagging in Slovene language is practical with proper use of Natural Language Toolkit and \textit{nltk-trainer}. Time spent in seconds for tagging different numbers of words, show that Brill and Trigram ran at the same speed, while Naive Bayes ran drasticly slower. There are no major differences in tagging accuracy. However Brill and the Naive Bayes are about 2\% better than Trigram.
\par
Besides these three taggers we built, others can be built without any bigger modification. All that is needed is to set some arguments when calling the script \texttt{train\_tagger.py}. The main reason we built only these three is computing power. These three are the only one we could built and evaluate in some resonable time.
\par
We also contacted the NLTK developers group and will arrange our taggers to become the part of NLTK.

@@

\begin{thebibliography}{99}
\bibitem{nltk} \url{http://www.nltk.org/}
\bibitem{JOS} \url{http://nl.ijs.si/jos/}
\bibitem{nltk-trainer} \url{https://github.com/japerk/nltk-trainer}
\bibitem{MULTEXT-East} \url{http://nl.ijs.si/ME/V4/msd/html/msd-sl.html}
\bibitem{NLTKBOOK} \url{http://www.nltk.org/book}
\bibitem{xml_tags} \url{http://nl.ijs.si/ssj/schema/tei_ssj_doc.pdf}
\bibitem{tokenizer} \url{http://docs.huihoo.com/nltk/0.9.5/guides/tokenize.html}
\end{thebibliography}
\end{document}

>>>>

We must exmplain, that here only the part of whole MSD is printed. For example the whole MSD for \textit{Lep} shown in table \ref{tab:MSDLep}.

\begin{table}[h]
\begin{center}
\begin{tabular}{c|c}
\multicolumn{2}{c}{pridevnik}\\\hline\hline
vrsta & splošni \\
spol & moški \\
število & ednina \\
sklon & imenovalnik \\
živost & 0 \\
vid & 0 \\
oblika & 0 \\
oseba & 0 \\
nikalnost & 0 \\
stopnja & nedoločeno \\
določnost & ne \\
število\_svojine & 0 \\
spol\_svojine & 0 \\
naslonskost & 0 \\
zapis & 0 \\
\end{tabular}
\end{center}
\caption{Whole MSD for \textit{Lep} in Slovene}
\label{tab:MSDLep}
\end{table}

The first output was creted with function \textit{prettyPrintWithDescription}, full output can be created with \textit{prettyPrintWithFullDescription}. Both function are defined in \texttt{example.py}
\par
But there is an issue. Because Slovenian language contain letters with caron, the output will have unicode symbols for them. In the previous examples, the output is already corrected. Solution is described in the following section.

\subsection{Encoding problems} %BANIČ
Printing the result of our example is not optimal. Because we want users to know what these characters are, we implemented a solution. To resolve this, we must decode every word to \textit{UTF - 8} (because words we tag are in UTF - 8). If a user wants to print the result in "pretty" form, he should call function \textit{prettyPrint()} or \textit{prettyPrintByLine()} (both are implemented in \texttt{example.py}).

@@

\section{Conclusion} %BANIČ
To conclude our paper, tagging in Slovene language is practical with proper use of Natural Language Toolkit and \textit{nltk-trainer}. Time spent in seconds for tagging different numbers of words, show that Brill and Trigram ran at the same speed, while Naive Bayes ran drasticly slower. There are no major differences in tagging accuracy. However Brill and the Naive Bayes are about 2\% better than Trigram.
\par
Besides these three taggers we built, others can be built without any bigger modification. All that is needed is to set some arguments when calling the script \texttt{train\_tagger.py}. The main reason we built only these three is computing power. These three are the only one we could built and evaluate in some resonable time.
\par
Our whole project is publicly available on github \cite{CODE}. We also contacted the NLTK developers group and will arrange our taggers to become the part of NLTK.

@@

\begin{thebibliography}{99}
\bibitem{nltk} \url{http://www.nltk.org/}
\bibitem{JOS} \url{http://nl.ijs.si/jos/}
\bibitem{nltk-trainer} \url{https://github.com/japerk/nltk-trainer}
\bibitem{MULTEXT-East} \url{http://nl.ijs.si/ME/V4/msd/html/msd-sl.html}
\bibitem{NLTKBOOK} \url{http://www.nltk.org/book}
\bibitem{xml_tags} \url{http://nl.ijs.si/ssj/schema/tei_ssj_doc.pdf}
\bibitem{tokenizer} \url{http://docs.huihoo.com/nltk/0.9.5/guides/tokenize.html}
\bibitem{CODE} \url{https://github.com/nikicc/slovene-nltk-tagger}
\end{thebibliography}
\end{document}
